---
title: "Statistical Inference report 2"
author: "C. Deramond"
date: "Sunday, December 21, 2014"
output: pdf_document
---

For this report we'll review results for analysis on the ToothGrowth dataset.


## 1. Load the ToothGrowth data and perform some basic exploratory data analyses 

```{r}
library(datasets)
library(reshape2)
library(dplyr, warn.conflicts = F)
library(ggplot2)
data <- ToothGrowth
```

Initially let's see if there's any difference in dosage:

```{r, echo=FALSE}
qplot(dose, data = data, bin= .3, 
      facets= .~supp, 
      title = "Distribution per supplement of dosage")
```

as there is no noticeable difference in dosage, we move on to compare distributions of lenghts:

```{r, echo=FALSE}
qplot(len, data = data, geom = "density", col = supp, tittle = "Distribution per supplement of lenght") 
```
so it's not a bad hypothesis that there is significant difference between lenght gained in average between supplements

## 2. Provide a basic summary of the data.

We'll summarise the data for the whole dataset:

```{r}
data %>%
        melt() %>%
        group_by(variable) %>%
        summarise(Mean = mean(value), SD = sd(value), Variance = var(value), n = n() ) %>%
        print    
```

Analogous process, so to divide by the supplement:

```{r, echo=FALSE}
data %>%
        melt() %>%
        group_by(variable, supp ) %>%
        summarise(Mean = mean(value), SD = sd(value), Variance = var(value), n = n() ) %>%
        print  
```


## 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)


## 4. State your conclusions and the assumptions needed for your conclusions. 