---
title: "Statistical Inference report 2"
author: "C. Deramond"
output: pdf_document
---

For this report we'll review results for analysis on the ToothGrowth dataset.


## 1. Load the ToothGrowth data and perform some basic exploratory data analyses 

```{r}
library(datasets)
library(reshape2)
library(dplyr, warn.conflicts = F)
library(ggplot2)
data <- ToothGrowth
```

Initially let's see if there's any difference in dosage:

```{r, echo=FALSE}
qplot(dose, data = data, bin= .3, 
      facets= .~supp, 
      title = "Distribution per supplement of dosage")
```

as there is no noticeable difference in dosage, we move on to compare distributions of lenghts:

```{r, echo=FALSE}
qplot(len, data = data, geom = "density", col = supp, tittle = "Distribution per supplement of lenght") 
```
at first glance: it might not be a bad hypothesis that there is significant difference in average lenght gained between supplements.

## 2. Provide a basic summary of the data.

We'll summarise the data for the whole dataset:

```{r}
data %>%
        melt() %>%
        group_by(variable) %>%
        summarise(Mean = mean(value), SD = sd(value), Variance = var(value), n = n() ) %>%
        print    
```

Analogous process, so to divide by the supplement:

```{r, echo=FALSE}
data %>%
        melt() %>%
        group_by(variable, supp ) %>%
        summarise(Mean = mean(value), SD = sd(value), Variance = var(value), n = n() ) %>%
        print  
```


## 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)

As shown by the summary table, there's not much sense in comparing dosage among supplements, as they are exactly equal. What might be of interest is to review whether is there a difference among the lengths obtained by different supplements:

* note the t-confidence interval for the lenght non discriminating by supplement: `r t.test(x=data$len)$conf.int`

so, it's safe to assume there is no difference among supplements as the mean is contained in the interval.

Digging deeper: we'll try the 2 sample non-paired t-test for $H_0$: $\mu_1=\mu_2$ under different variances:

```{r}
# for safety we assume different variances for non-paired t-test
t.test(len ~ supp, paired = F, data = data, var.equal = F)
```

so we fail to reject null hypothesis.

## 4. State your conclusions and the assumptions needed for your conclusions. 

* There's no significant difference in length gained between the 2 supplements.
* Note that the only assumptions we are taking are the standard considerations for a Welch's test.